Modelling the Lymphatic Metastatic Progression Pathways of OPSCC from Multi-Institutional Datasets

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

\autoref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well. Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data. The separation between involvement prevalences of early and late T-category tumors is also reproduced well by the model. This is remarkable because the model introduces only a single parameter to describe the differences between early and late T-category for all involvement patterns. This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors can explain more severe lymphatic involvement.

@@

Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$. The model evidence for these three graphs is reported in \autoref{table:evidence}. The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V. Interestingly, the evidence gain of this ``winning graph'' is roughly the sum of the gains seen in the two candidates where only one of these connections was added, respectively. This indicates that the two additional parameters are largely independent of each other and manage to describe different aspects of the data.

\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III and from level III to IV is removed. The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation. The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.

>>>>

\autoref{fig:bg_prevalences} shows that this minimal graph is already capable of describing the most important parts of the observed data very well. Notably, the model is not only accurate in its predictions, it also correctly estimates the variance stemming from the limited amount of data. The separation between involvement prevalences of early and late T-category tumors is also reproduced well by the model. This is remarkable because the model introduces only a single parameter to describe the differences between early and advanced T-category for all involvement patterns. This shows that expecting later diagnosis times, on average, for patients with advanced T-category tumors can explain more severe lymphatic involvement.

@@

Based on these results, we consider three additional candidates for the optimal graph that combine the added arc from \gls{lnl} I to II with the arc(s) $\text{III} \rightarrow \text{V}$ and/or $\text{IV} \rightarrow \text{V}$. The model evidence for these three graphs is also reported in \autoref{table:evidence}. The best performing graph with decisive evidence over the base graph turned out to be the one which combines the arcs from \gls{lnl} I to II and IV to V. Interestingly, the evidence gain of this ``winning graph'' is roughly the sum of the gains seen in the two candidates where only one of these connections was added, respectively. This indicates that the two additional parameters are largely independent of each other and manage to describe different aspects of the data.

\autoref{table:evidence} additionally shows the evidence of graphs in which the arc from level II to III or from level III to IV is removed. The low model evidence for these graphs confirms the importance of these connections and is consistent with the anatomical motivation. The connection from level III to IV is crucial for describing the observation that metastases in level IV are extremely rare without simultaneous involvement of the upstream level III.
